\documentclass[11pt]{article}

\usepackage{fullpage,times}%charter}
\usepackage{color}

\usepackage{tikz}
\usetikzlibrary{arrows.meta}

%% macros
\newcommand{\ax}[1]{\texttt{AX}(#1)}
\newcommand{\ex}[1]{\texttt{EX}(#1)}
\newcommand{\af}[1]{\texttt{AF}(#1)}
\newcommand{\ef}[1]{\texttt{EF}(#1)}
\newcommand{\ag}[1]{\texttt{AG}(#1)}
\newcommand{\eg}[1]{\texttt{EG}(#1)}
\newcommand{\au}[2]{\texttt{A}(#1\ \texttt{U}\ #2)}
\newcommand{\eu}[2]{\texttt{E}(#1\ \texttt{U}\ #2)}
\newcommand{\sem}[1]{[\!\![#1]\!\!]}

\newcommand{\sol}[1]{{\color{blue}#1}}

\begin{document}

\hrule
\smallskip

\noindent
You can review the latex source for this assignment-file to
learn and use latex to prepare your homework submission. You will see
the use of macros (to write uniformly formatted text), different
text-styles (emphasized, bold-font), different environments (figures,
enumerations).

It is not required that you use exactly this latex source to prepare
your submission. 
\smallskip
\hrule


\begin{center}
{\Large\bf Homework 1 (CTL): ComS/CprE/SE 412, ComS 512}

\medskip

Due-date: Feb 7 at 11:59PM.

\medskip


\end{center}

\noindent
\textbf{
Submit online on Canvas two files: the source file in latex format and
the pdf file generated from latex. Name your files:
$\langle\mbox{your-net-id}\rangle$-hw1.$\langle\mbox{tex/pdf}\rangle$.
}

\hrule
\noindent
\smallskip

\emph{ Homework must be individual's original work. Collaborations and
  discussions of any form with any students or other faculty members
  or soliciting solutions on online forums are not allowed. Please
  review the academic dishonesty policy on our syllabus. If you have
  any questions/doubts/concerns, post your questions/doubts/concerns
  on Piazza and ask TA/Instructor.}

\smallskip
\hrule

\begin{enumerate}

\item
Consider the following Kripke structure, with $p\in L(s_0) \cap L(s_2)$ and
$q\in L(s_2)$.
\begin{center}
    \begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
    \node (s0) at (0,0) {$s_0$};
    \node (s1) at (2,0) {$s_1$};
    \node (s2) at (4,0) {$s_2$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black,very thick}]
    \path [->] (s0) edge[bend left=30] (s1);
    \path [->] (s1) edge[bend left=30] (s0);
    \path [->] (s1) edge[bend left=30] (s2);
    \path [->] (s2) edge[bend left=30] (s1);
    \path [->] (s2) edge[loop above] (s2);
\end{scope}
\end{tikzpicture}
\end{center}
Identify the set of states that satisfy each of the following:
\begin{enumerate}
\item $\ex{q}$

  
\item $\ax{p}$
  
\item $\ax{q}$

\item $\ag{p}$
  
\item $\eg{p}$

\item $\af{p}$

\item $\ag{\ex{p}}$
  
\item $\ag{\af{p}}$
  
\end{enumerate}
\hfill(16 pts)

\item Express the following statements as CTL formula:\hfill (4+4+6 pts)
\begin{enumerate}

\item
Along all paths \textbf{withdraw-money} is never true after
\textbf{invalid-login}.

\textbf{Answer}: AG (\textbf{invalid-login} $\Rightarrow$ AXAG  \textbf{withdraw-money})

\item
Along all execution sequences of an elevator behavior, if the elevator
door is \textbf{open} then the door remains \textbf{open} until a
\textbf{request-to-move} is sent to the elevator.

\textbf{Answer}: AG ((\textbf{open} $\Rightarrow$ AG  \textbf{open}) U \textbf{request-to-move})

\item 
Whenever proposition \textbf{request-site-update} is true in a state,
it is followed in zero or more steps by a state where proposition
\textbf{updating-site} is true, which in turn is followed in one or
more steps by a state where \textbf{update-complete} is true.

\textbf{Answer}: AG ( \textbf{request-site-update} $\Rightarrow $AF \textbf{updating-site} ) $\Rightarrow$ AXAF \textbf{update-complete})


\end{enumerate}


\item To disprove that two CTL formula are equivalent, you are
  required to draw a Kripke structure and identify a state in that
  structure, which satisfies one of the formula and does not satisfy
  the other. For instance, in order to disprove that
  $\ex{p}\ \land\ \ex{q}$ and $\ex{p\ \land\ q}$ are equivalent, we can
  draw the following the Kripke structure:

  \begin{center}
  \begin{tabular}{cc}
  \begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
    \node (s0) at (0,0) {$s_0$};
    \node (s1) at (-1,-1) {$s_1$};
    \node (s2) at (1,-1) {$s_2$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black,very thick}]
    \path [->] (s0) edge (s1);
    \path [->] (s0) edge  (s2);
    \path [->] (s1) edge[loop below] (s1);
    \path [->] (s2) edge[loop below] (s2);
\end{scope}
  \end{tikzpicture}
  &
  %%
  \begin{tabular}{l}
    $L(s_1) = \{p\}$, $L(s_2) = \{q\}$
    \end{tabular}
  \end{tabular}
\end{center}
We present the labels for the relevant states in the Kripke structure
and state that $s_0$ satisfies $\ex{p}\ \land\ \ex{q}$ and $s_0$ does
not satisfy $\ex{p \land q}$. This is because there exists a path from
$s_0$ where in the very next state $s_1$ the proposition $p$ is true;
thus conforming that $s_0$ satisfies $\ex{p}$ (similarly, one can justify
the satisfiability of $\ex{q}$ at state $s_0$). On the other hand,
there exists no path from $s_0$, where in the very next state both $p$
and $q$ are satisfied. 
\smallskip

  
Disprove that the two CTL formula $\af{p \land q}$ and
$\af{p} \land \af{q}$ are equivalent.  \hfill (5 pts)

\item
To prove that one formula (say, $\varphi_1$) is ``stronger'' than
another (say, $\varphi_2$), you need to prove that whenever in any
state in any Kripke structure $\varphi_1$ holds, $\varphi_2$ holds in
that state as well. In other words, $\varphi_1$ is ``stronger'' than
$\varphi_2$ if and only if $\varphi_1\Rightarrow\varphi_2$ is a
tautology (always evaluates to true). For instance, to prove that
$\ax{p}$ is stronger than $\ex{p}$ we can write:
%% math environment starts
\[
\begin{array}{rcl}
  \forall s. s\in \sem{\ax{p}} & \Leftrightarrow & \forall \pi\in Path(s): \pi[1] \in \sem{p} \\
                               & \Rightarrow &     \exists \pi\in Path(s): \pi[1] \in \sem{p} \\
                               & \Rightarrow & s \in \sem{\ex{p}}
\end{array}
\]
%% math environment ends
%%
To disprove that $\ex{p}$ is stronger that $\ax{p}$, you will draw a
Kripke structure and present a state in the Kripke structure that
satisfies $\ex{p}$ but does not satisify $\ax{p}$ (see the Kripke
structure example in the previous problem).

Prove or disprove $\ag{p\ \land\ q}$ is stronger than $\ag{p\ \Rightarrow\ \af{q}}$.


\hfill (5 pts)

\item \textbf{Extra Credit for 412; Required problem for 512.}
  We will define a new operator $A\circ$
  as follows. A state $s$ satisfies $\texttt{A}(\varphi_1\ \circ\ \varphi_2)$
  if and only if for all paths starting from $s$ at least one of the
  following holds
\begin{itemize}
\item there exists a state where $\varphi_2$ is
  satisfied and before that $\varphi_1$ holds in all states in the
  path
\item all states in the path satisfy $\varphi_1$ and not
  $\varphi_2$
\end{itemize}
Prove or disprove that 
\begin{enumerate}
\item $A(p\ \circ\ \mbox{false})$ can be expressed
  in CTL.

\item $A(\mbox{false}\ \circ\ p)$ can be expressed in CTL.

\end{enumerate}
\hfill(10pts)
\end{enumerate}

\end{document}

